DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-13 paper/background.tex
Changes to paper

The use of neural networks for end-to-end control of autonomous
car was first demonstrated in late 1980s~\cite{Pomerleau1989},
using a small 3-layer fully connected neural network; and subsequently
in a DARPA Autonomous Vehicle (DAVE) project in early
2000s~\cite{LeCun:04}, using a 6 layer convolutional neural network
(CNN); and most recently in NVIDIA's DAVE-2
project~\cite{Bojarski2016}, using 9 layer CNN. In all these projects,
the neural network models take raw image pixels as input and directly
produce steering control commands bypassing all intermediary steps and
hand-written rules in conventional robotics control approach.  
NVIDIA's latest effort reports that their trained CNN
autonomously controls their modified cars on public roads without human
intervention successfully~\cite{Bojarski2016}.

Using deep neural networks involves two distinct
phases~\cite{NVIDIA2015}. The first
phase is \emph{training} during which the weights of the network is
incrementally updated by backpropagating errors it sees from the
training examples. Once the network is trained---i.e., the weights of
the network minimize errors in the training examples---the next phase
is \emph{inferencing} during which unseen data is fed to the network
as input to produce predicted output (e.g., predicted image
classification). In general, the training phase is most computationally
intensive and requires high throughput, which is generally not
available on embedded platforms. The inferencing phase, on the
other hand, is relatively less computationally intensive and latency becomes
as important, if not more, as computational throughput because many
use-cases have strict real-time requirements (e.g., search query
latency).

@@

\subsection{Embedded Computing Platforms for Real-Time Inferencing}
Real-time embedded systems, such as an autonomous vehicle, present
unique challenges for deep learning as the computing platforms of such
systems must satisfy two often conflicting goals~\cite{Otterness2017}:
%% Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
%% in large part to the increased computing performance,
%% which afforded researchers to train and use ever deeper neural networks with
%% high accuracy.
%% For practical applications, the computer platform in an
%% autonomous vehicle must satisfy two often conflicting goals:
The platform must provide 
enough computing capacity for real-time processing of computationally
expensive AI workloads (deep neural networks);
The platform must also satisfy various
constraints such as cost, size, weight, and power consumption limits.

Accelerating AI workloads, especially inferencing
operations, has received a lot of attention from academia and industry
in recent years as applications of deep learning are broadening to
areas real-time embedded systems such as autonomous vehicles. These
efforts include development of various heterogeneous architecture
based system-on-a-chips (SOCs) that may include multiple cores, GPU,
DSP, FPGA and neural network optimized ASIC hardware~\cite{Jouppi2017}.
Consolidating multiple tasks on SoCs with lots of shared hardware
resources while guaranteeing real-time performance is also an active
research area, which is orthogonal to improving raw
performance. Consolidation is necessary for efficiency but unmanaged 
interference can nullify the benefits of consolidation~\cite{Kim2016}.
Because of these reasons, finding a good computing platform is a
non-trivial task, one that requires deep understanding of the
workloads and the hardware platform being utilized.

The primary objectives of this study are (1) to understand the
necessary computing performance for applying AI technology based
robotics systems, and (2) what kind of computing architecture and
runtime supports are most appropriate for such workload. Toward to
achieve these goals, we implement a low-cost autonomous car platform
as a case-study and systematically conduct experiments, as we will 
describe in the subsequent sections.

>>>>

The use of neural networks for end-to-end control of autonomous
car was first demonstrated in late 1980s~\cite{Pomerleau1989},
using a small 3-layer fully connected neural network; and subsequently
in a DARPA Autonomous Vehicle (DAVE) project in early
2000s~\cite{LeCun:04}, using a 6 layer convolutional neural network
(CNN); and most recently in NVIDIA's DAVE-2
project~\cite{Bojarski2016}, using a 9 layer CNN. In all of these projects,
the neural network models take raw image pixels as input and directly
produce steering control commands, bypassing all intermediary steps and
hand-written rules used in conventional robotics control approach.  
NVIDIA's latest effort reports that their trained CNN
autonomously controls their modified cars on public roads without human
intervention~\cite{Bojarski2016}.

Using deep neural networks involves two distinct
phases~\cite{NVIDIA2015}. The first
phase is \emph{training} during which the weights of the network are
incrementally updated by backpropagating errors it sees from the
training examples. Once the network is trained---i.e., the weights of
the network minimize errors in the training examples---the next phase
is \emph{inferencing}, during which unseen data is fed to the network
as input to produce predicted output (e.g., predicted image
classification). In general, the training phase is most computationally
intensive and requires high throughput, which is generally not
available on embedded platforms. The inferencing phase, on the
other hand, is relatively less computationally intensive and latency becomes
as important, if not moreso, as computational throughput, because many
use-cases have strict real-time requirements (e.g., search query
latency).

@@

\subsection{Embedded Computing Platforms for Real-Time Inferencing}
Real-time embedded systems, such as an autonomous vehicle, present
unique challenges for deep learning, as the computing platforms of such
systems must satisfy two often conflicting goals~\cite{Otterness2017}:
%% Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
%% in large part to the increased computing performance,
%% which afforded researchers to train and use ever deeper neural networks with
%% high accuracy.
%% For practical applications, the computer platform in an
%% autonomous vehicle must satisfy two often conflicting goals:
The platform must provide 
enough computing capacity for real-time processing of computationally
expensive AI workloads (deep neural networks);
The platform must also satisfy various
constraints such as cost, size, weight, and power consumption limits.

Accelerating AI workloads, especially inferencing
operations, has received a lot of attention from academia and industry
in recent years as applications of deep learning are broadening to
areas of real-time embedded systems such as autonomous vehicles. These
efforts include the development of various heterogeneous architecture-based 
system-on-a-chips (SOCs) that may include multiple cores, GPU,
DSP, FPGA and neural network optimized ASIC hardware~\cite{Jouppi2017}.
Consolidating multiple tasks on SoCs with lots of shared hardware
resources while guaranteeing real-time performance is also an active
research area, which is orthogonal to improving raw
performance. Consolidation is necessary for efficiency but unmanaged 
interference can nullify the benefits of consolidation~\cite{Kim2016}.
For these reasons, finding a good computing platform is a
non-trivial task, one that requires deep understanding of the
workloads and the hardware platform being utilized.

The primary objectives of this study are to understand (1) the
necessary computing performance for applying AI technology-based
robotics systems, and (2) what kind of computing architecture and
runtime supports are most appropriate for such workload. To
achieve these goals, we implement a low-cost autonomous car platform
as a case-study and systematically conduct experiments, which we will 
describe in the subsequent sections.

### 2017-10-13 paper/comparison.tex
Changes to paper

In this section, we compare three computing platforms---the Raspberry
Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and NVIDIA
Jetson
TX2~\footnote{http://www.nvidia.com/object/embedded-systems-dev-kits-modules.html}---from
the point of view of supporting vision-based end-to-end deep learning
based autonomous vehicles. 
Table~\ref{tbl:platforms} shows architectural features of the three
platforms~\footnote{The GPUs of the Raspberry Pi 3 and Intel
  Up are not used in evaluation due to the lack of software (TensorFlow)
support. Also, the two Denver cores in Tegra TX2 are not used in
evaluation due to TensorFlow issues.}.

@@

The UP Board and TX2 also perform much better when multiple DNN models
are co-scheduled. Figure~\ref{fig:sys_model} shows the results of the
multi-model co-scheduling experiment. Once again, they can comfortably
satisfy 20 Hz real-time performance for all of the co-scheduled DNN control
loops, and in case of TX 2 (GPU), 100 Hz real-time control is still
feasible. Given that the GPU must be shared among the co-scheduled DNN
models, the result suggests that TX2's GPU has sufficient capacity to
accomodate multiple instances of the DNN model we tested.

%% to complete control loops in a fraction of the time when compared to 
%% the Pi. Furthermore, both of the platforms satisfied the 50 ms WCET 
%% in all experiments, while the Raspberry Pi 3 was unable to do so. As 
%% such, the UP Board and TX2 display greater potential for the 
%% simultaneous execution of multiple models during self-driving 
%% operation, while the Raspberry Pi 3 would struggle to do so.

@@

Another intersting observation is that TX2 (GPU) also suffers
considerable execution time increase (2.5X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architectue based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

%% experienced a change in their inferencing times, but the increases 
%% were not as drastic as is seen by the Raspberry Pi 3. in the worst 
%% case, both the UP Board and TX2 (CPU only and GPU) produced times 
%% that were over two times as large, whereas the Raspberry Pi 3 output 
%% times that were around 10 times greater. The Intel UP Board was 
%% successful in completing inferencing operations in under 50 ms when 
%% benchmarks were run on a maximum of 2 cores, and the TX2 was 
%% successful in all experiments irrespective of the computing resources 
%% utilized. Comparetively, both performed better than the Raspberry Pi 
%% 3 which failed to meet its deadlines when any benchmark was 
%% introduced. As such, it can be concluded that the Intel UP Board 
%% would be more capable of running computationally heavy processes 
%% during autonomous driving, and that the Pi wouldn't be able to do the 
%% same.

@@

%% the comparison of the real-time capabilities of three embedded 
%% computing platforms, we found that the Raspberry Pi 3 performed the 
%% worst. When compared to the Intel UP Board and the NVIDIA Tegra TX2, 
%% control loops on the Pi were shown to take noticably longer to 
%% complete, regardless of the number of cores utilized or the number of 
%% neural network models run simultaneoulsy. The difference was even 
%% more noticeable in the the addition of computationally heavy 
%% synthetic benchmarks that had a much more dire effect on the Pi. In 
%% the worst case, an average control loop on the Pi took an abnormally 
%% long amount of time to finish. Compared to the TX2 with GPU support, 
%% those times were 34.8X times greater. From this, we find that the UP 
%% Board and Tegra TX2 are superior platforms for real-time CNN 
%% operations. However, due to the Pi's ability to satisfy the 50 ms 
%% WCET in some experiments, we conclude that the Raspberry Pi 3 is 
%% feasible for autonomous research.

>>>>

In this section, we compare three computing platforms---the Raspberry
Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and NVIDIA
Jetson
TX2~\footnote{http://www.nvidia.com/object/embedded-systems-dev-kits-modules.html}---from
the point of view of supporting end-to-end deep learning
based autonomous vehicles. 
Table~\ref{tbl:platforms} shows architectural features of the three
platforms~\footnote{The GPUs of the Raspberry Pi 3 and Intel
  Up are not used in evaluation due to the lack of software (TensorFlow)
support. Also, the two Denver cores in Tegra TX2 are not used in
evaluation due to TensorFlow issues.}.

@@

The Intel UP board and TX2 also perform much better when multiple DNN models
are co-scheduled. Figure~\ref{fig:sys_model} shows the results of the
multi-model co-scheduling experiment. Once again, they can comfortably
satisfy 20 Hz real-time performance for all of the co-scheduled DNN control
loops, and in case of TX 2 (GPU), 100 Hz real-time control is still
feasible. Given that the GPU must be shared among the co-scheduled DNN
models, the result suggests that TX2's GPU has sufficient capacity to
accomodate multiple instances of the DNN model we tested.

@@

Another interesting observation is that TX2 (GPU) also suffers
considerable execution time increase (2.5X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architectue based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

@@

### 2017-10-13 paper/conclusion.tex
Changes to paper

%% We evaluated the real time efficacy of our platform and found
%% that it is capable of performing inferencing operations within a WCET
%% of 50 ms, or 20 fps, when a single model is run on more than one
%% core.

As future work, we plan to apply shared resource management
techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
evaluate their impact to the real-time performance of the system. We
also plan to improve the prediction accuracy by feeding more data and
upgrading the RC car hardware platform to enable more precise steering
angle control.

%% Further studies should be conducted on the interference that
%% occurs between cores to allow for improved performance in utilizing
%% multiple models for autonomous vehicles. With minimal further
%% research, the DeepPicar can become a valudable, reliable and
%% inexpensive tool in the continued research of autonomous vehicles.

>>>>

As future work, we plan to apply shared resource management
techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
evaluate their impact on the real-time performance of the system. We
also plan to improve the prediction accuracy by feeding more data and
upgrading the RC car hardware platform to enable more precise steering
angle control.
